Typing the Outputs

### 2015-10-09 sematics/boundary/boundary.tex
some old changes to boundary type stuff

\newcommand {\colfunsym} {:>}
\newcommand {\colfun} [3] {{#1}\colfunsym{#2}~\myatsign~{#3}}

@@

\begin{framed}
\noindent\textbf{First-Order Recursive Functions:}
\begin{mathpar}
\infertypeswor [\to I]		{\Gamma \vdash {\fix ux{e}} \colfunsym {A \to B}}	{\types [\Gamma, u \colfunsym {A \to B},\col x A] {e} B} 							\and
\infertypeswor [\to E]		{\types {\app {f}{e}} {B}}							{\Gamma \vdash {f} \colfunsym {A \to B} & \ty {e} A} 					\and
\infertypeswor [\fut E]		{\types {\letfin u {f} {e}} C}						{\Gamma \vdash {f} \colfunsym {A \to B} 
																				&\types [\Gamma, u \colfunsym {A \to B}] e C } 					\and
\end{mathpar}
\end{framed}
The important thing to note here is that functions have only two forms: direct declaration or reference to an assumed variable.

@@

As in the monostage system, first-stage first-order functions don't participate in the value/computation system.
Second-stage functions do, however, since we need to mark whether they contain first stage code.
\begin{framed}
\noindent\textbf{First-Order Recursive Functions:}
\begin{mathpar}
\infertypeswor [\to I]		{\Gamma \vdash \colfun {\fix ux{e}} {A \to B} w}	{\ty [,\colfun u {A \to B} w,\col x A] {e} B} 							\and
%\infertypeswor [fix]		{\ty {\fix uxe} {A \to B}}							{\ty [,\col f {A \to B},\col x A] e B} 										\and
\infertypeswor [\to E]		{\ty {\app {f}{e}} {B}}								{\Gamma \vdash \colfun f {A \to B} w & \ty e A} 					\and
\infertypeswor [\fut E]		{\ty {\letfin u f e} B}								{\Gamma \vdash \colfun f {A \to B} w 
																				&\ty [,\colfun u {A \to B} w] e B } 					\and
\infer {\Gamma \vdash \fix uxe \valw}											{\Gamma, u \valw, x \valw \vdash e \valw}				\and
\infer {\Gamma \vdash \fix uxe \compt}											{\Gamma, u \valt, x \valt \vdash e \compt}				\and
\infer {\Gamma \vdash \app{f}{e} \compw}										{\Gamma \vdash e \compw}								\and
\infer {\Gamma \vdash \app{f}{e} \valt}											{\Gamma \vdash e \valt}									\and
\infer {\Gamma \vdash \letfin{u}f{e} \compo}									{\Gamma \vdash e \compo}								\and
\infer {\Gamma \vdash \letfin{u}f{e} \compt}									{\Gamma, u \compt \vdash e \compt}
\end{mathpar}
\end{framed}

@@

\section{First Order Recursive Functions}
\label{sec:firstorder}

@@

\subsection {Functions}

@@

>>>>

\newcommand {\colfunsym} {:>}
\newcommand {\colfun} [3] {{#1}\colfunsym{#2}~\myatsign~{#3}}
\newcommand {\unfunwrap}[1] {\lfloor{#1}\rfloor}
\newcommand {\funwrap}[1] {\lceil{#1}\rceil}

@@

\begin{framed}
\noindent\textbf{First-Order Recursive Functions:}
\begin{mathpar}
\infertypeswor [\to I]		{\Gamma \vdash {\fix ux{e}} \colfunsym {A \to B}}	{\types [\Gamma, u \colfunsym {A \to B}, x : A] e B} 							\and
\infertypeswor [\to E]		{\types {\app {f}{e}} {B}}							{\Gamma \vdash {f} \colfunsym {A \to B} & \types e A} 					\and
\infertypeswor [\fut E]		{\types {\letfin u {f} {e}} C}						{\Gamma \vdash {f} \colfunsym {A \to B} 
																				&\types [\Gamma, u \colfunsym {A \to B}] e C } 					\and
\end{mathpar}
\end{framed}
The important thing to note here is that functions have only two forms: direct declaration or reference to an assumed variable.

\subsection{Higher-Order Second-Class Functions}

We can extend this formalism to include functions which are higher-order but still second-class.

\begin{framed}
\noindent\textbf{Higher-Order Second-Class Functions:}
\begin{mathpar}
\infertypeswor [\to I]		{\Gamma \vdash \funwrap e \colfunsym \overline A}	{\types e A} 													\and
\infertypeswor [\to I]		{\types {\unfunwrap f} A}							{\Gamma \vdash f \colfunsym \overline A} 						\and
\infertypeswor [\to I]		{\Gamma \vdash {\lam uf} \colfunsym {A \to B}}		{\Gamma, u \colfunsym A \vdash f \colfunsym B} 					\and
\infertypeswor [\to E]		{\types {\app {f_1}{f_2}} {B}}						{\Gamma \vdash {f_1} \colfunsym {A \to B} 
																				&\Gamma \vdash {f_2} \colfunsym A} 								\and
\infertypeswor [\fut E]		{\types {\letfin u {f} {e}} C}						{\Gamma \vdash {f} \colfunsym {A \to B} 
																				&\types [\Gamma, u \colfunsym {A \to B}] e C } 					\and
\end{mathpar}
\end{framed}

@@

As in the monostage system, first-stage first-order functions don't participate in the value/computation system.
Second-stage functions do, however, since we need to mark whether they contain first stage code.
\begin{framed}
\noindent\textbf{First-Order Recursive Functions:}
\begin{mathpar}
\infertypeswor [\to I]		{\Gamma \vdash \colfun {\fix ux{e}} {A \to B} w}	{\ty [,\colfun u {A \to B} w,\col x A] {e} B} 							\and
%\infertypeswor [fix]		{\ty {\fix uxe} {A \to B}}							{\ty [,\col f {A \to B},\col x A] e B} 										\and
\infertypeswor [\to E]		{\ty {\app {f}{e}} {B}}								{\Gamma \vdash \colfun f {A \to B} w & \ty e A} 					\and
\infertypeswor [\fut E]		{\ty {\letfin u f e} C}								{\Gamma \vdash \colfun f {A \to B} w 
																				&\ty [,\colfun u {A \to B} w] e C } 					\and
\infer {\Gamma \vdash \fix uxe \valw}											{\Gamma, u \valw, x \valw \vdash e \valw}				\and
\infer {\Gamma \vdash \fix uxe \compt}											{\Gamma, u \valt, x \valt \vdash e \compt}				\and
\infer {\Gamma \vdash \app{f}{e} \compw}										{\Gamma \vdash e \compw}								\and
\infer {\Gamma \vdash \app{f}{e} \valt}											{\Gamma \vdash e \valt}									\and
\infer {\Gamma \vdash \letfin{u}f{e} \compo}									{\Gamma \vdash e \compo}								\and
\infer {\Gamma \vdash \letfin{u}f{e} \compt}									{\Gamma, u \compt \vdash e \compt}
\end{mathpar}
\end{framed}

\begin{framed}
\noindent\textbf{Higher-Order Second-Class Functions:}
\begin{mathpar}
\infertypeswor [\to I]		{\Gamma\vdash \colfun{\funwrap e}{\overline A} w}	{\ty e A} 													\and
\infertypeswor [\to I]		{\ty {\unfunwrap f} A}								{\Gamma\vdash \colfun f{\overline A} w} 						\\
\infertypeswor [\to I]		{\Gamma \vdash \colfun {\lam xe} {A \to B} w}		{\Gamma, \colfun x A w \vdash \colfun e B w} 			\and
%\infertypeswor [fix]		{\ty {\fix uxe} {A \to B}}							{\ty [,\col f {A \to B},\col x A] e B} 										\and
\infertypeswor [\to E]		{\Gamma \vdash \colfun {\app {f_1}{f_2}} {B} w}		{\Gamma \vdash \colfun {f_1} {A \to B} w 
																				&\Gamma \vdash \colfun {f_2} A w} 						\and
\infertypeswor [\fut E]		{\ty {\letfin u f e} B}								{\Gamma \vdash \colfun f A w 
																				&\ty [,\colfun u A w] e B} 								\\
\infer {\Gamma \vdash \funwrap e \valo}											{\Gamma \vdash e \valo}									\and
\infer {\Gamma \vdash \unfunwrap f \valo}										{\Gamma \vdash f \valo}									\\
\infer {\Gamma \vdash \lam xe \valo}											{\Gamma, x \valo \vdash e \compo}						\and
\infer {\Gamma \vdash \lam xe \valt}											{\Gamma, x \valt \vdash e \valt}						\and
\infer {\Gamma \vdash \lam xe \compt}											{\Gamma, x \valt \vdash e \compt}						\and
\infer {\Gamma \vdash \app{f}{e} \compw}										{\Gamma \vdash e \compw}								\and
\infer {\Gamma \vdash \app{f}{e} \valt}											{\Gamma \vdash e \valt}									\and
\infer {\Gamma \vdash \letfin{u}f{e} \compo}									{\Gamma \vdash e \compo}								\and
\infer {\Gamma \vdash \letfin{u}f{e} \compt}									{\Gamma, u \valt \vdash e \compt}
\end{mathpar}
\end{framed}

@@

\section{Second-Class Functions}
\label{sec:firstorder}

@@

\subsection {First-Order Functions}

@@

\subsection {Higher-Order Functions}

Now we consider an extension to higher-order, but still second-class, functions.
The main difference is that now, after the $\ibtsym$ we can have a type constructor, 
which is either an actual type, a nullary-tuple constructor, or a function from one type constructor to another.

\begin{framed}
\noindent\textbf{Higher-Order Function Boundary Types:}
\begin{mathpar}
\infer {\ibtyo {\funwrap e} {\overline A} {()}}   {\cdot} \and
\infer {\btyo {\unfunwrap e} {\overline A} \rmunit}   {\cdot} \and
\infer {\ibtyo {\lam uf} {A \to B} {\lambda \alpha.\sigma}}
	{\ibtyo [\Gamma, \alpha \wf, \ibtj u A {\alpha}] f B {\sigma}} \and
\infer {\ibtyo {\app {f_1}{f_2}} {B} {(\sigma_1~\sigma_2)}} 
	{\ibtyo {f_1} {A \to B} {\sigma_1} 
	&\ibtyo {f_2} A {\sigma_2}} \and
\infer {\btyo {\letin u f e} C \tau}
	{\ibtyo {f} A {\sigma} 
	&\btyo [\Gamma, \ibtj u A \sigma] {e} C {\tau}}
% \infer {\btyt {\fix uxe} {A} {\tau}} 
% 	{\btyt {e} A {\tau}} \and
% \infer {\btyt {\app {f}{e}} {A} {\tau}} 
% 	{\btyt {e} A {\tau}} \and
% \infer {\btyw {\letin x {e_1} {e_2}} {B} {\tau_1 \times \tau_2}} 
% 	{\btyw {e_1} A {\tau_1} 
% 	&\btyw {e_2} B {\tau_2}}
\end{mathpar}
\end{framed}
